Meta-learning in movement prediction problem of aperiodic time-series data

### 2024-10-04 doc/paper/paper.tex
draft 3

\documentclass[aps,prb,groupedaddress,twocolumn,showpacs,dvipdfmx,superscriptaddress,pdftex]{revtex4-2}
\usepackage[whole]{bxcjkjatype}
\usepackage{amsmath}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\usepackage{amssymb}
% \usepackage{bm}
% \usepackage{dcolumn}
% \usepackage{bm}
\usepackage{url}
% \usepackage{here}
% \usepackage{ulem}
% \usepackage[caption=false]{subfig}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[usenames,dvipsnames]{color}

\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{rotating}
% \usepackage{makecell}
\usepackage{multirow}

@@

% tiếng Việt
\usepackage[T5]{fontenc}
\usepackage[utf8]{inputenc}
\DeclareTextSymbolDefault{\DH}{T1}
%-------------------------------------
\newcommand{\tadd}[1]{\textcolor{red}{#1}}
\newcommand{\trem}[1]{{\color{red}\sout{#1}}}
\newcommand{\tcom}[1]{\textcolor{red}{\textbf{TI: #1}}}
\newcommand{\citehere}{\textcolor{red}{\textbf{(citation)}}}
\newcommand{\x}{\times}
%-------------------------------------
% \newcommand{\ghost}[1]{[#1]}
\newcommand{\ghost}[1]{}
%----------------------------------------
\def\red{\color{red}}
\def\blue{\color{blue}}
\def\green{\color{green}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title{
    Meta-learning in movement prediction problem of aperiodic time-series data
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Bao-Long Nguyen}
\email{mwklng2309@icloud.com}
\affiliation{School of Information Science, JAIST, Ishikawa, Japan}
%
\author{Kenta Hongo}
\email{kenta_hongo@mac.com}
\affiliation{Research Center for Advanced Computing Infrastructure, JAIST, Ishikawa, Japan}
%
\author{Ryo Maezono}
\email{rmaezono@mac.com}
\affiliation{School of Information Science, JAIST, Ishikawa, Japan}
%
\author{Tom Ichibha}
\email{ichibha@icloud.com}
\affiliation{School of Information Science, JAIST, Ishikawa, Japan}
%
\date{\today}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

@@

For the first challenge, ensemble models \citep{sadeghi2021combined, zafeiriou2020intraday, ali2020complete} are often used to mitigate the effects of variance variation. Ensemble models provide a holistic and multi-perspective view based on sub-models, thereby helping the overall model adapt to strong variance changes. However, under the rigid synthesis of traditional ensemble models, the results obtained on non-periodic datasets are very poor.

@@

To overcome the second challenge, most studies use features extracted from Long Short-Term Memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial Neural Network (\verb|ANN|), and Convolution Neural Network (\verb|CNN|)\cite{lecun1989handwritten}. Specifically, in 2022, 20\% of all publications related to financial index prediction used \verb|LSTM|, 20\% used \verb|ANN|, and 6\% used \verb|CNN|\cite{ayitey2023forex}. This is completely understandable because \verb|LSTM| and \verb|CNN| have proven their ability to extract spatial and temporal features effectively. However, when faced with aperiodic data with only 2600 samples, after 40 training steps, \verb|LSTM| is still underfitting, while \verb|CNN| is overfitting after only 5 training steps.

@@

Regarding the third challenge, \cite{fama1970efficient} hypothesizes that different time-series datasets of the same domain at the same time reflect the impact of external factors. For example, studies \citep{overreactioncontrarian, mech1993portfolio} show the dependence between the financial ratios of a given company and the ratios of other companies. This further strengthens the hypothesis in study \cite{fama1970efficient}. Therefore, by exploiting data sources with the same domain, we can capture information reflecting external factors.

@@

In summary, our main contributions include the following: First, we present a novel \textbf{feature combination technique} that extracts and combines features using both \verb|LSTM| and \verb|CNN| models. Second, we uncover hidden long-term dependencies, demonstrating experimentally that aperiodic time-series data not only rely on external factors but also exhibit dependencies with past period, which we call \textbf{hidden long-term dependencies}. Third, we introduce an \textbf{efficient model parameter aggregation approach} by employing ML algorithm in place of traditional ensemble models for combining parameters from multiple machine learning models. Finally, we validate our method through experiments on exchange rate datasets and show its effectiveness in comparison to the SOTA model, \verb|NHITS|.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related work}
\label{sec.relatedWork}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Model-agnostic Meta-learning (MAML)}

@@

Meta-learning (ML) algorithms, typically MAML \cite{finn2017model} are known for their ability to train a highly general, adaptive model on new datasets with a limited amount of data and a small number of training steps \citep{hospedales2021meta, vettoruzzo2024advances}. With this ability, ML is widely used in tasks that require the model's ability to adapt to the data (e.g. personalization of learning models \citep{chen2018federated, fallah2020personalized,nguyen2022meta}, domain adaptation in online learning \citep{hu2023meta, khoee2024domain}).

@@

Hybrid ensemble models have been widely used in time-series processing problems and have been experimentally proven to be more accurate than standard time-series models because they can synthesize the strengths of many sub-models \cite{ayitey2023forex}. However, current ensemble model synthesis forms are still very rigid because they can only synthesize based on the final results (voting mechanism of bagging models) and near-final results (for stacking models). From the perspective of ensemble models, the equation \ref{eq:outer_opt} can be considered an effective method of synthesizing sub-models, which helps to take advantage of the feature extraction capabilities of each model. In other words, the synthesized model can extract features at a deeper level, significantly improving the prediction ability compared to traditional ensemble models.

@@

% Trong đó, \verb|FullyConnected| là các lớp multi-layer perception (\verb|MLP|) xếp chồng với hàm kích hoạt phi tuyến. $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ là các hệ số nội suy forecast và backcast, được dùng để tổng hợp các giá trị đầu ra của block $l$ bằng hàm nội suy $g(\cdot)$. Đầu ra của block $l$ là giá trị forecast $\mathbf{\hat{y}}_{t+1:t+H, l}$ và giá trị backcast $\mathbf{\hat{y}}_{t-L:t, l}$. Input của block $l+1$ được tính theo phương trình \ref{eq:input_l1}.

Accordingly, \verb|FullyConnected| are stacked multi-layer perception (\verb|MLP|) layers with nonlinear activation functions. $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ are the forecast and backcast interpolation coefficients, which are used to aggregate the output values of block $l$ using the interpolation function $g(\cdot)$. The output of block $l$ is the forecast value $\mathbf{\hat{y}}_{t+1:t+H, l}$ and the backcast value $\mathbf{\hat{y}}_{t-L:t, l}$. The input of block $l+1$ is calculated according to the equation \ref{eq:input_l1}.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{img/CHF_NZD.pdf}
    \caption{Exchange rate (close price) between Swiss franc and New Zealand dollar by day (2014-2024).}
    \label{fig:fx}
\end{figure}

@@

Inspired by the \cite{vo2017multi} study, we propose to combine the features extracted from \verb|LSTM| and \verb|CNN| networks. Specifically, we pass each element in the vector $\mathbf{x}_{t-L:t}$ through a \verb|MLP| layer whose output'dimension is larger than the one of $\vec x_i, i\in[t-L, t]$ to decompose it into smaller features $\vec x'_i$. These features are then passed through \verb|LSTM| and \verb|CNN| networks to extract long-term temporal dependencies ($\mathbf{h}_{LSTM}$) and local temporal features ($\mathbf{h}_{CNN}$), respectively. To exploit the long-term temporal constraints, we use \verb|BidirectionalLSTM| to extract from both sides of $\mathbf{x}_{t-L:t}$. The entire feature extraction process is summarized as follows:

@@

The \verb|LSTM| network maintains cell-state values ​​to selectively store long-term dependencies. This is very suitable for solving time-series data problems. On the other hand, future values ​​often depend heavily on recent historical values. We propose to use the \verb|CNN| network to emphasize local features, thereby directing part of the model's attention to certain time points. Therefore, the proposed method can not only remember long-term features but also highlight short-term features.

@@

We use \verb|MAML| to train and aggregate the weights of the models at the tasks. As mentioned in \ref{sec.relatedWork}, parameter optimization in the ML approach is to solve the two equations \ref{eq:inner_opt} and \ref{eq:outer_opt} using optimization methods on the support and query data. Specifically, the optimization process includes many global steps (outer optimization), performed on all tasks participating in training. Each global step includes many local steps (inner optimization) performed on each individual task. At global step $r$, the $e$th local optimization process at the support set of task $t$ occurs as follows:

@@

FX in particular and financial indices in general are typical data types for aperiodic time-series data. Therefore, we choose this type of data to test the model. Specifically, we configure two datasets using FX data. The \verb|USD/JPY| dataset consists of only data of the USD/JPY currency pair, divided into 60 time-sequenced subsets of equal size. The data is sampled hourly from 2000 to 2024, including the attributes of open, low, high, and close price. The \verb|multi-fx| dataset consists of 60 currency pairs made up of 18 countries: Australia, Canada, Switzerland, Denmark, EU, United Kingdom, Hong Kong, Iceland, Japan, Norway, New Zealand, Singapore, Sweden, Turkey, United States, Mexico, China, South Africa. The data has similar attributes to \verb|USD/JPY| and sampled daily from 2014 to 2024. The number of data samples of these two datasets are similar and approximately $156000$ samples.

@@

The \verb|multi-fx| dataset is used to extract and aggregate information about outliers (i.e. market, economic, political, etc.) that are believed to influence the outcome of a given financial index \citep{overreactioncontrarian,fama1970efficient,mech1993portfolio}. The \verb|USD/JPY| dataset is used to test our hypothesis that future data implicitly depend on certain points in the past and that efficient past feature aggregation is needed to reveal these dependencies.

@@

The study uses macro accuracy, macro precision, macro recall, and macro F1 score to evaluate the models. Accordingly, during the inference phase, the model will run on each task to calculate the metrics of each task. Then, the average of the metrics of the tasks is calculated to obtain the final result.

@@

We compare the proposed method with \verb|NHITS| using the above metrics on \verb|USD/JPY| and \verb|multi-fx| datasets. We also test different feature extractors (\verb|LSTM|, \verb|CNN|) for the proposed method to demonstrate the deep data understanding capabilities of \verb|LSTM+CNN|. The overall data is structured into training sets, validation sets, and testing sets with a ratio of 6:2:2 for training, fine-tuning and testing model, respectively.

@@

For \verb|NHITS| model, we split the data as usual according to the above predetermined ratio. For the proposed algorithm, because the meta-training and meta-testing processes require dividing the data into small tasks, and allowing the model to adapt on the support set of each task, we split the data into 60 tasks (see table \ref{tab:stat_} for statistical detail). In each task, the support set accounts for 20\% with the purpose of letting the model adapt to the data, the query set accounts for 80\% to check the model's compatibility. We use 30 tasks for meta-training, 15 tasks for meta-validation, and 15 tasks for meta-testing. With this division, we ensure the fairness that the ML model is trained with the same amount of data as \verb|NHITS|. After the data configuration, we perform experiment as in appendix \ref{app:our_experiment} and \ref{app:nhits_experiment}.

\begin{table}
    \centering
    \caption{Statistics on USD/JPY and multi-fx.}
    \label{tab:stat_}
    \begin{tabular}{ccccccc}
    \toprule
    \multirow{2}{*}{Dataset} & \multirow{2}{*}{\#task} & \multirow{2}{*}{\#samples} & \multicolumn{4}{c}{\#samples/task}  \\
    \cline{4-7}
                             &                         &                            & min & mean  & std   & max             \\
    \hline
    \Verb|USD/JPY|           & 60                      & 150,175                     & 2,502 & 2,502.92 & 0.28 & 2,503           \\
    \Verb|multi-fx|          & 60                      & 154,440                     & 2,467 & 2,575 & 58.61   & 2748           \\
    \bottomrule
    \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Result \& Discussion}
\label{sec.results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Bảng \ref{tab:rs} so sánh kết quả giữa mô hình đề xuất và mô hình \verb|NHITS| trên các tập dữ liệu kể trên. Phương pháp của chúng tôi đạt mức hội tụ cao hơn hẳn so với mô hình \verb|NHITS| thể hiện trên tất cả các metrics. Cụ thể, độ chính xác cải thiện từ 5\% đến 11\% so với \verb|NHITS| trên dữ liệu \verb|USD/JPY| và \verb|multi-fx|. Các metrics còn lại cũng đều tăng từ 5\% đến 20\%. Sau đây, chúng tôi phân tích kết quả này dựa trên hai yếu tố: (1) - Khả năng rút trích đặc trưng; (2) - Khả năng tổng hợp mô hình.

Table \ref{tab:rs} compares the results between the proposed model and the \verb|NHITS| model on the above datasets. Our method achieves a much higher convergence rate than the \verb|NHITS| model on all metrics. Specifically, the accuracy improves from 5\% to 11\% compared to \verb|NHITS| on \verb|USD/JPY| and \verb|multi-fx| data. The remaining metrics also increase from 5\% to 20\%. Next, we analyze this result based on two factors: (1) - Feature extraction ability; (2) - Model synthesis ability.

\begin{table*}
    \centering
    \caption{Classification results (\%) of NHITS and our method using USD/JPY and multi-fx datasets. Best results per metrics are boldfaced.}
    \label{tab:rs}
    \begin{tabular}{c|c|cccc}
    \toprule
    \multicolumn{1}{c}{}                          &                                    & $accuracy$            & $precision$           & $recall$              & $F1$                    \\
    \hline
    \multirow{3}{*}{\Verb|USD/JPY|}  & \Verb|NHITS|          & 53.29                 & 52.38                 & 52.23                 & 51.87                   \\
                                                & Ours(\Verb|LSTM|)     & $57.12±0.08$ & $58.32±2.72$ & $57.44±2.62$ & $56.87±3.56$    \\
                                                & Ours(\Verb|LSTM+CNN|) & $\mathbf{58.05±0.01}$ & $\mathbf{59.57±2.53}$ & $\mathbf{58.31±2.57}$ & $\mathbf{56.6±3.74}$    \\
    \hline
    \multirow{3}{*}{\Verb|multi-fx|} & \Verb|NHITS|          & 51.12                 & 51.22                 & 51.16                 & 50.54                   \\
                                                & Ours(\Verb|LSTM|)     & $61.23±0.06$          & $70.52±9.17$          & $69.51±9.31$          & $68.9±10.0$             \\
                                                & Ours(\Verb|LSTM+CNN|) & $\mathbf{62.39±0.06}$ & $\mathbf{71.11±9.34}$ & $\mathbf{70.14±9.6}$  & $\mathbf{69.21±11.07}$  \\
    \bottomrule
    \end{tabular}
\end{table*}

>>>>

\documentclass[aps,prb,groupedaddress,twocolumn,showpacs,dvipdfmx,superscriptaddress,pdftex]{revtex4-2}
\usepackage[whole]{bxcjkjatype}
\usepackage{amsmath}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\usepackage{amssymb}
\usepackage{url}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[usenames,dvipsnames]{color}

\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{rotating}
\usepackage{multirow}

@@

% tiếng Việt
\usepackage[T5]{fontenc}
\usepackage[utf8]{inputenc}
\DeclareTextSymbolDefault{\DH}{T1}
%-------------------------------------
\newcommand{\tadd}[1]{\textcolor{red}{#1}}
\newcommand{\trem}[1]{{\color{red}\sout{#1}}}
\newcommand{\tcom}[1]{\textcolor{red}{\textbf{TI: #1}}}
\newcommand{\citehere}{\textcolor{red}{\textbf{(citation)}}}
\newcommand{\x}{\times}
%-------------------------------------
% \newcommand{\ghost}[1]{[#1]}
\newcommand{\ghost}[1]{}
%----------------------------------------
\def\red{\color{red}}
\def\blue{\color{blue}}
\def\green{\color{green}}

% \usepackage{makecell}
% \usepackage{bm}
% \usepackage{dcolumn}
% \usepackage{here}
% \usepackage{ulem}
% \usepackage[caption=false]{subfig}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title{
    Meta-learning in movement prediction problem of aperiodic time-series data
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Bao-Long Nguyen}
\email{mwklng2309@icloud.com}
\affiliation{School of Information Science, JAIST, Ishikawa, Japan}
%
\author{Kenta Hongo}
\email{kenta_hongo@mac.com}
\affiliation{Research Center for Advanced Computing Infrastructure, JAIST, Ishikawa, Japan}
%
\author{Ryo Maezono}
\email{rmaezono@mac.com}
\affiliation{School of Information Science, JAIST, Ishikawa, Japan}
%
\author{Tom Ichibha}
\email{ichibha@icloud.com}
\affiliation{School of Information Science, JAIST, Ishikawa, Japan}
%
\date{\today}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

@@

For the variance variation challenge, ensemble models \citep{sadeghi2021combined, zafeiriou2020intraday, ali2020complete} are often used to mitigate the effects of variance variation. Ensemble models provide a holistic and multi-perspective view based on sub-models, thereby helping the overall model adapt to strong variance changes. However, under the rigid synthesis of traditional ensemble models, the results obtained on non-periodic datasets are very poor.

@@

To overcome the pattern variation challenge, most studies use features extracted from Long Short-Term Memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial Neural Network (\verb|ANN|), and Convolution Neural Network (\verb|CNN|) \cite{lecun1989handwritten}. Specifically, in 2022, 20\% of all publications related to financial index prediction used \verb|LSTM|, 20\% used \verb|ANN|, and 6\% used \verb|CNN| \cite{ayitey2023forex}. This is completely understandable because \verb|LSTM| and \verb|CNN| have proven their ability to extract spatial and temporal features effectively. However, when faced with aperiodic data with only 2600 samples, after 40 training steps, \verb|LSTM| is still underfitting, while \verb|CNN| is overfitting after only 5 training steps.

@@

Regarding the external influences challenge, \cite{fama1970efficient} hypothesizes that different time-series datasets of the same domain at the same time reflect the impact of external factors. For example, studies \citep{overreactioncontrarian, mech1993portfolio} show the dependence between the financial ratios of a given company and the ratios of other companies. This further strengthens the hypothesis in study \cite{fama1970efficient}. Therefore, by exploiting data sources with the same domain, we can capture information reflecting external factors.

@@

In summary, our main contributions include the following: First, we present a novel \textbf{feature combination technique} for aperiodic time-series data that extracts and combines features using both \verb|LSTM| and \verb|CNN| models. Second, we uncover hidden long-term dependencies, demonstrating experimentally that aperiodic time-series data not only rely on external factors but also exhibit dependencies with past period, which we call \textbf{hidden long-term dependencies}. Third, we introduce an \textbf{efficient model parameter aggregation approach} by employing ML algorithm in place of traditional ensemble models for combining parameters from multiple machine learning models. Finally, we validate our method through experiments on exchange rate datasets and show its effectiveness in comparison to the SOTA model, \verb|NHITS|.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related works}
\label{sec.relatedWork}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Optimization-based Meta-learning}

@@

Optimization-based Meta-learning (ML) algorithms, typically MAML \cite{finn2017model} are known for their ability to train a highly general, adaptive model on new datasets with a limited amount of data and a small number of training steps \citep{hospedales2021meta, vettoruzzo2024advances}. With this ability, ML is widely used in tasks that require the model's ability to adapt to the data (e.g. personalization of learning models \citep{chen2018federated, fallah2020personalized,nguyen2022meta}, domain adaptation in online learning \citep{hu2023meta, khoee2024domain}).

@@

Hybrid ensemble models have been widely used in time-series processing problems and have been experimentally proven to be more accurate than standard time-series models because they can synthesize the strengths of many sub-models \cite{ayitey2023forex}. However, current ensemble model synthesis forms are still very rigid because they can only synthesize based on the final results (e.g. voting mechanism of bagging models) and near-final results (e.g. stacking models). From the perspective of ensemble models, the equation \ref{eq:outer_opt} can be considered an effective method of synthesizing sub-models, which helps to take advantage of the feature extraction capabilities of each model. In other words, the synthesized model can extract features at a deeper level, significantly improving the prediction ability compared to traditional ensemble models.

@@

% Trong đó, \verb|FullyConnected| là các lớp multi-layer perception (\verb|MLP|) xếp chồng với hàm kích hoạt phi tuyến. $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ là các hệ số nội suy forecast và backcast, được dùng để tổng hợp các giá trị đầu ra của block $l$ bằng hàm nội suy $g(\cdot)$. Số chiều của các hệ số nội suy trong mỗi stack được quy định bởi expressiveness ratio $r_l$: $\left| \theta_l \right| =  \left \lceil r_l H \right \rceil$. Thông thường, expressiveness ratio sẽ rất nhỏ ở các stack đầu tiên và tăng dần về cuối. Theo đó, các stack có thể mô phỏng lại các tần số từ thấp đến cao. Ngoài ra, bằng việc sử dụng hàm nội suy, \verb|NHITS| không cần quá nhiều phần cứng để huấn luyện mạng neural trong trường hợp H lớn.

Accordingly, \verb|Pooling| compresses the original data into $\mathbf{y}_{t-L:t, l}^{(p)}$. \verb|FullyConnected| contains many perceptron layers with nonlinear activation functions. $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ are the forecast and backcast interpolation coefficients, which are used to aggregate the output values of block $l$ using the interpolation function $g(\cdot)$. The dimensionality of the interpolation coefficients in each stack is determined by the expressiveness ratio: $\left| \theta_l \right| = \left \lceil r_l H \right \rceil$. Typically, the expressiveness ratio will be very small in the first stacks and increase towards the end. Accordingly, the stacks can simulate frequencies from low to high. In addition, by using $g(\cdot)$, \verb|NHITS| does not require too much hardware to train the neural network in the case of large $H$.

% Đầu ra của block $l$ là giá trị forecast $\mathbf{\hat{y}}_{t+1:t+H, l}$ và giá trị backcast $\mathbf{\hat{y}}_{t-L:t, l}$. Input của block $l+1$ được tính theo phương trình \ref{eq:input_l1}.

The output of block $l$ is the forecast value $\mathbf{\hat{y}}_{t+1:t+H, l}$ and the backcast value $\mathbf{\hat{y}}_{t-L:t, l}$. The input of block $l+1$ is calculated according to the equation \ref{eq:input_l1}.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{img/USD_JPY.pdf}
    \caption{Exchange rate (close price) between US dollar and Japanese yen by day (2014-2024).}
    \label{fig:fx}
\end{figure}

@@

Inspired by the \cite{vo2017multi} study, we propose to combine the features extracted from \verb|LSTM| and \verb|CNN| networks. Specifically, we pass each element in the vector $\mathbf{x}_{t-L:t}$ through \verb|FullyConnected| layers whose output'dimension is larger than the one of $\vec x_i, i\in[t-L, t]$ to decompose it into smaller features $\vec x'_i$. These features are then passed through \verb|LSTM| and \verb|CNN| networks to extract long-term temporal dependencies ($\mathbf{h}_{LSTM}$) and local temporal features ($\mathbf{h}_{CNN}$), respectively. To exploit the long-term temporal constraints, we use \verb|BidirectionalLSTM| to extract from both sides of $\mathbf{x}_{t-L:t}$. The entire feature extraction process is summarized as follows:

@@

The \verb|LSTM| network maintains cell-state values to selectively store long-term dependencies. This is very suitable for solving time-series data problems. On the other hand, future values often depend heavily on recent historical values. We propose to use the \verb|CNN| network to emphasize local features, thereby directing part of the model's attention to certain time points. Therefore, the proposed method can not only remember long-term features but also highlight short-term features.

@@

We use \verb|MAML| to train and aggregate the weights of the models at the tasks. As mentioned in \ref{sec.relatedWork}, parameter optimization in the ML approach is to solve the two equations \ref{eq:inner_opt} and \ref{eq:outer_opt} using optimization methods on the support and query data. Specifically, the optimization process includes many global steps (outer optimization), performed on all tasks participating in training. Each global step includes many local steps (inner optimization) performed on each individual task. At global step $r$-th, the $e$-th local optimization process at the support set of task $t$ occurs as follows:

@@

FX in particular and financial indices in general are typical data types for aperiodic time-series data. Therefore, we choose this type of data to test the model. Specifically, we configure two datasets using FX data. \verb|USD/JPY| dataset consists of only data of the exchange rate between US dollar and Japanese yen, is divided into 60 time-sequenced subsets of equal size. The data is sampled hourly from 2000 to 2024, including the attributes of open, low, high, and close price. \verb|multi-fx| dataset consists of 60 currency pairs made of 18 countries: Australia, Canada, Switzerland, Denmark, EU, United Kingdom, Hong Kong, Iceland, Japan, Norway, New Zealand, Singapore, Sweden, Turkey, United States, Mexico, China, South Africa. The data has similar attributes to \verb|USD/JPY| and sampled daily from 2014 to 2024.

@@

Our goal is to use \verb|multi-fx| dataset to extract and aggregate information about outliers (i.e. market, economic, political, etc.) that are believed to influence the outcome of a given financial index \citep{overreactioncontrarian,fama1970efficient,mech1993portfolio}. Meanwhile \verb|USD/JPY| dataset is used to test our hypothesis that future data implicitly depend on certain points in the past and that efficient past feature aggregation is needed to reveal these dependencies.

% Ngoài ra, chúng tôi sử dụng thêm 2 bộ dữ liệu: Electricity Transformer Temperature (\verb|ETT-m2|) và Weather (\verb|WTH|). Tập dữ liệu \verb|ETT-m2| bao gồm 7 trường dữ liệu, đo đạc các thông số của máy biến áp tại một tỉnh của Trung Quốc sau mỗi 15 phút trong khoảng thời gian từ July 2016 to July 2018. Tập dữ liệu \verb|WTH| bao gồm 12 trường dữ liệu, ghi nhận các thông số của thời tiết tại Weather Station of the Max Planck Biogeochemistry Institute in Jena, Germany sau mỗi 10 phút trong năm 2020. Đây là hai tập dữ liệu thể hiện tính chu kỳ rất mạnh mà \verb|NHITS| đã rất thành công trong việc dự đoán. Thực nghiệm trên chúng giúp so sánh toàn diện hơn về khả năng của phương pháp đề xuất đối với \verb|NHITS|.

In addition, we used two periodic datasets: Electricity Transformer Temperature (\verb|ETT-m2|) \cite{zhou2021informer} and Weather (\verb|WTH|) \cite{Kolle}. \verb|ETT-m2| dataset consists of 7 data fields, measuring the parameters of a transformer in a province of China every 15 minutes from July 2016 to July 2018. \verb|WTH| dataset consists of 12 data fields, recording the weather parameters at the Weather Station of the Max Planck Biogeochemistry Institute in Jena, Germany every 10 minutes in 2020. These two datasets exhibit very strong periodicity, which \verb|NHITS| has been very successful in predicting. Experiments on them provide a more comprehensive comparison of the proposed method's capabilities against \verb|NHITS|.

@@

The study uses macro accuracy, macro precision, macro recall, and macro F1 score to evaluate the models. Accordingly, during the inference phase, the model will run on all tasks to calculate the metrics of each one. Then, the average of the metrics of the tasks is calculated to obtain the final result.

@@

We compare the proposed method with \verb|NHITS| using the above metrics and dataset. For each dataset, each attribute will be predicted for future trends based on past data. The final result is calculated by averaging the metrics across the attributes. We also test different feature extractors (\verb|LSTM|, \verb|CNN|) for the proposed method to demonstrate the deep data understanding capabilities of \verb|LSTM+CNN|. The overall data is structured into training sets, validation sets, and testing sets with a ratio of 6:2:2 for training, fine-tuning and testing model, respectively.

@@

For \verb|NHITS| model, we split the data as usual according to the above predetermined ratio. For the proposed algorithm, because the meta-training and meta-testing processes require dividing the data into small tasks, and allowing the model to adapt on the support set of each task, we split the data into 40-60 tasks (see table \ref{tab:stat_} for statistical detail). In each task, the support set accounts for 20\% with the purpose of letting the model adapt to the data, the query set accounts for 80\% to check the model's compatibility. We use 50\% of tasks for meta-training, 25\% for meta-validation, and 25\% for meta-testing. With this division, we ensure the fairness that the ML model is trained with the same amount of data as \verb|NHITS|. After the data configuration, we perform experiment as in appendix \ref{app:our_experiment} and \ref{app:nhits_experiment}.

\begin{table}
    \centering
    \caption{Statistics on USD/JPY and multi-fx.}
    \label{tab:stat_}
    \begin{tabular}{ccccccc}
    \toprule
    \multirow{2}{*}{Dataset} & \multirow{2}{*}{\#task} & \multirow{2}{*}{\#samples} & \multicolumn{4}{c}{\#samples/task}  \\
    \cline{4-7}
                             &                         &                            & min & mean  & std   & max             \\
    \hline
    \Verb|USD/JPY|           & 60                      & 150,175                     & 2,502 & 2,502.92 & 0.28 & 2,503           \\
    \Verb|multi-fx|          & 60                      & 154,440                     & 2,467 & 2,575 & 58.61   & 2748           \\
    \Verb|ETT-m2|            & 48                      & 69,680                      & 1,451 & 1,451.67 & 0.47   & 1,452           \\
    \Verb|WTH|               & 40                      & 35,064                      & 876 & 876.6 & 0.49   & 877           \\
    \bottomrule
    \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Result \& Discussion}
\label{sec.results}
% TODO: phải viết lại đống nhận xét dựa trên kết quả mới
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Bảng \ref{tab:rs} so sánh kết quả giữa mô hình đề xuất và mô hình \verb|NHITS| trên các tập dữ liệu kể trên. Trên dữ liệu phi chu kỳ (\verb|multi-fx| and \verb|USD/JPY|), phương pháp của chúng tôi đạt mức hội tụ cao hơn hẳn so với mô hình \verb|NHITS| thể hiện trên tất cả các metrics. Cụ thể, độ chính xác cải thiện từ 5\% đến 11\% so với \verb|NHITS| trên dữ liệu \verb|USD/JPY| và \verb|multi-fx|. Các metrics còn lại cũng đều tăng từ 5\% đến 20\%. Trên dữ liệu có chu kỳ, phương pháp của chúng tôi ghi nhận các kết quả cao hơn hoặc xấp xỉ kết quả của \verb|NHITS|. Cụ thể, trên tập \verb|WTH|, phương pháp của chúng tôi cho giá trị cao hơn khoảng 1\% trên tất cả các độ đo, trên tập \verb|ETT-m2|, chúng tôi thu được độ chính xác cao hơn khoảng 1\% trong khi các metrics khác thấp hơn khoảng 3\%. Sau đây, chúng tôi phân tích kết quả này dựa trên hai yếu tố: (1) - Khả năng rút trích đặc trưng; (2) - Khả năng tổng hợp mô hình.

Table \ref{tab:rs} compares the results between the proposed model and the \verb|NHITS| model on the above datasets. On aperiodic data, our method achieves a much higher convergence rate than the \verb|NHITS| model on all metrics. Indeed, the accuracy improves from 5\% to 11\% compared to \verb|NHITS| on \verb|USD/JPY| and \verb|multi-fx| data. The remaining metrics also increase from 5\% to 20\%. On periodic data (\verb|ETT-m2| and \verb|WTH| dataset), our method obtains results that are higher than or close to those of \verb|NHITS|. Specifically, our results differ by 1-3\% on all metrics compared to the results of \verb|NHITS|. In the following subsections, we analyze these results on aperiodic datasets based on two factors: (1) - Feature extraction ability; (2) - Model synthesis ability; and relate to periodic datasets.
